\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-09 theory.tex
minor
edited out Marika (temporarily on her request)
CH typo check read through I think this is ready to send...

Figure~\ref{fig:Cl_xi} shows theoretical models for the shear power spectrum $P_\gamma(\ell)$ (equation~\ref{eqn:Pkappa}), and the shear correlation function (equation~\ref{eqn:xiGG}) for the different variations on the approximations listed in Table~\ref{tab:Tl_nu}.   The Standard and Spherical models agree at the few percent level above $\ell>10$ and below $\theta< 100$ arcmin.  The two outliers are the Baseline and Extended `Flat' Sky Limber cases (LF and ELF) where the \citet{kitching/etal:2016} `Flat' Sky approximation has been made with
\be
T_\ell^{\rm `Flat'} = \ell^4 / (\ell + 0.5)^4 = 1-{2\over \ell}+{5\over 2\ell^2} +{\cal O}(\ell^{-3}) \, . 
\ee    
We argue that applying the approximation, that $\ell \pm x \approx \ell$, where $x$ is small, only to the numerator of equation~\ref{eqn:Tl} is inconsistent.  By comparing the expansions of equations~\ref{eqn:Tl} and~\ref{eqn:Tlflat}, we see that the \citet{kitching/etal:2016} `Flat' Sky approximation effectively introduces a new term $2/\ell$ making it more deviant from the full spherical solution in comparison to the standard $T_\ell = 1$ case.    To our knowledge, this form of the flat-sky approximation has not been used in cosmic shear studies to date.

As shown in the lower panels of Figure~\ref{fig:Cl_xi}, the `standard flat' sky approximation used by recent cosmic shear surveys\footnote{The KSt approximation was used for pre-2014 CFHTLenS analysis and the ESt approximation was used for recent CFHTLenS and KiDS analyses}, where $T_\ell = 1$, recovers a very similar result to the spherical sky case.  Over angular scales used in the recent KiDS cosmic shear analysis (where the maximum angular scale $\theta<50.7$ arcmin is shown dashed), these curves differ by less than 1\%.  As such the standard flat-sky approximations that have been used will not impact the cosmological analyses of current surveys.

@@

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber-approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses, carried out many years ago, remained unpublished in peer reviewed journals as no significant deviation was found\footnote{See for example appendix M of the 2010 PhD thesis from Donghui Jeong (\url{http://www.personal.psu.edu/duj13/dissertation/djeong_diss_appM.pdf}) and Chapter 7 of 2012 PhD thesis from Nicolas Van de Rijt (\url{http://ipht.cea.fr/Docspht/articles/t12/080/public/these_vanderijt2012.pdf}).}.  One notable exception, however, is \citet{giannantonio/etal:2012} who find that for a cosmic shear power spectrum, the extended Limber approximation is consistent with the exact calculation for $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  They conclude that the Limber approximation works so well for cosmic shear owing to the wide redshift distribution of the lensed sources.  This is in comparison to the case of galaxy power spectra,  where narrower redshift bins result in a significant difference between the Limber and non-Limber cases \citep[see also][]{simon/2007}.   Given this body of work, we argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for the cosmic shear power spectrum, individual teams have also verified this result internally for their own cosmic shear analyses.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in many theoretical codes over the years.  Instead these codes have evolved to account for critical approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    Two exceptions to this are recent updates to CLASS \citep{blas/lesgourgues/tram:2011,audren/etal:2013} and CosmicFish \citep{raveri/etal:2016}.  On-going code comparisons have, however, revealed differences that are larger than the expected impact of Limber or non-Limber approximation.   We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} to draw their own conclusions about the impact of this effect.

>>>>

Figure~\ref{fig:Cl_xi} shows theoretical models for the shear power spectrum $P_\gamma(\ell)$ (equation~\ref{eqn:Pkappa}), and the shear correlation function (equation~\ref{eqn:xiGG}) for the different variations on the approximations listed in Table~\ref{tab:Tl_nu}.   The Standard and Spherical models agree at the few percent level above $\ell>20$ and below $\theta< 100$ arcmin.  The two outliers are the Baseline and Extended `Flat' Sky Limber cases (LF and ELF) where the \citet{kitching/etal:2016} `Flat' Sky approximation has been made with
\be
T_\ell^{\rm `Flat'} = \ell^4 / (\ell + 0.5)^4 = 1-{2\over \ell}+{5\over 2\ell^2} +{\cal O}(\ell^{-3}) \, . 
\label{eqn:Tlflat}
\ee    
We argue that applying the approximation, that $\ell \pm x \approx \ell$, where $x$ is small, only to the numerator of equation~\ref{eqn:Tl} is inconsistent.  By comparing the expansions of equations~\ref{eqn:Tl} and~\ref{eqn:Tlflat}, we see that the \citet{kitching/etal:2016} `Flat' Sky approximation effectively introduces a new term $2/\ell$ making it more deviant from the full spherical solution in comparison to the standard $T_\ell = 1$ case.    To our knowledge, this form of the flat-sky approximation has not been used in cosmic shear studies to date.

As shown in the lower panels of Figure~\ref{fig:Cl_xi}, the `standard flat' sky approximation used by recent cosmic shear surveys\footnote{The KSt approximation was used for pre-2014 CFHTLenS analysis and the ESt approximation was used for recent CFHTLenS and KiDS analyses.}, where $T_\ell = 1$, recovers a very similar result to the spherical sky case.  Over angular scales used in the recent KiDS cosmic shear analysis (where the maximum angular scale $\theta<50.7$ arcmin is shown dashed), these curves differ by less than 1\%.  As such the standard flat-sky approximations that have been used will not impact the cosmological analyses of current surveys.

@@

 \subsection{Cosmic shear without the Limber Approximation}
In collecting our comments on \citet{kitching/etal:2016} we contacted colleagues who had previously tested the impact of the Limber approximation on cosmic shear studies by comparing a Limber-approximated cosmic shear power spectrum with an exact calculation.  The majority of these analyses remained unpublished in peer reviewed journals as no significant deviation was found\footnote{See for example appendix M of the 2010 PhD thesis from Donghui Jeong (\url{http://www.personal.psu.edu/duj13/dissertation/djeong_diss_appM.pdf}) and Chapter 7 of 2012 PhD thesis from Nicolas Van de Rijt (\url{http://ipht.cea.fr/Docspht/articles/t12/080/public/these_vanderijt2012.pdf}).}.  One notable exception, however, is \citet{giannantonio/etal:2012} who find that for a cosmic shear power spectrum, the extended Limber approximation is consistent with the exact calculation for $\ell>5$, with any differences well within cosmic-variance errors for a Euclid-like survey\footnote{\citet{kitching/etal:2016} comment on this result, but indicate that it is in error owing to the application of a fixed low-$\ell$ limit.  \citet{giannantonio/etal:2012} however state that this limit is only applied in a positive curvature case.}.  They conclude that the Limber approximation works well for cosmic shear owing to the wide redshift distribution of the lensed sources.  This is in contrast to the case of galaxy power spectra,  where narrower redshift bins result in a significant difference between the Limber and non-Limber cases \citep[see also][]{simon/2007}.   Given this body of work, we argue that the assertion by \citet{kitching/etal:2016} that the majority of cosmic shear data analyses to date have made an `axiomatic assumption (i.e. an unquestioned and untested assumption at the beginning of the analysis)'  is unwarranted.  Not only is there published work that has questioned and tested the Limber approximation for the cosmic shear power spectrum, individual teams have also verified this result internally for their own cosmic shear analyses.

Given the insignificant effect of using the Limber approximation found by various groups, the ability to calculate an exact non-Limber solution has unfortunately not been maintained in many theoretical codes over the years.  Instead these codes have evolved to account for critical approximations such as the impact of baryon feedback on the matter power spectrum and non-zero neutrino masses \citep[see for example][]{joudaki/etal:2016, mead/etal:2016}.    Two exceptions to this are recent updates to CLASS \citep{blas/lesgourgues/tram:2011,audren/etal:2013} and CosmicFish \citep{raveri/etal:2016}.  On-going code comparisons have, however, revealed differences between the calculated theoretical models that are larger than the expected impact of Limber or non-Limber approximation.   We are therefore unable to readily determine non-Limber theoretical calculations for the shear-correlation function to directly compare with the results of \citet{kitching/etal:2016}.   Instead we refer the reader to Figure 3 of \citet{giannantonio/etal:2012} to draw their own conclusions about the impact of this effect.
